---
title: "Paper_comparison"
output: github_document
date: "2026-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries. If needed, use BiocManager to install them
```{r loading libraries, echo=FALSE, include=FALSE}
library("phyloseq")
library("ggplot2")
library("vegan")
library("readr")
library("tidyverse")
library("RColorBrewer")
```

```{r package versions, echo=FALSE}
packageDescription("phyloseq", fields = c("Package", "Version"))
packageDescription("ggplot2", fields = c("Package", "Version"))
packageDescription("vegan", fields = c("Package", "Version"))
packageDescription("readr", fields = c("Package", "Version"))
packageDescription("tidyverse", fields = c("Package", "Version"))
packageDescription("RColorBrewer", fields = c("Package", "Version"))
```

# Importing files, renaming ASVs and creating phyloseq object - AUER 
```{r importing files auer}
seqtab.nochim_auer <- readr::read_csv("Auer_seqtab.nochim_16S.csv")
seqtab.nochim_auer <- as.data.frame(seqtab.nochim_auer)
rownames(seqtab.nochim_auer) <- seqtab.nochim_auer[, 1]
seqtab.nochim_auer <- seqtab.nochim_auer[, -1]
colnames(seqtab.nochim_auer) = paste0("ASV", 1:ncol(seqtab.nochim_auer)) 

taxa_auer <- read.csv("Auer_taxa_16S.csv", row.names = 1)
row.names(taxa_auer) = colnames(seqtab.nochim_auer)
taxa_auer <- as.matrix(taxa_auer)

metadata_auer <- read.csv("Metadata_AUER.csv", stringsAsFactors = TRUE, row.names = 1)

ps_auer <- phyloseq(
  otu_table(seqtab.nochim_auer, taxa_are_rows=FALSE),
  sample_data(metadata_auer),
  tax_table(taxa_auer))

ps_auer
```
# Importing files, renaming ASVs and creating phyloseq object - DENG_2017 
```{r importing files deng7}
seqtab.nochim_DENG7 <- readr::read_csv("Deng_seqtab.nochim_16S.csv")
seqtab.nochim_DENG7 <- as.data.frame(seqtab.nochim_DENG7)
rownames(seqtab.nochim_DENG7) <- seqtab.nochim_DENG7[, 1]
seqtab.nochim_DENG7 <- seqtab.nochim_DENG7[, -1]
colnames(seqtab.nochim_DENG7) = paste0("ASV", 1:ncol(seqtab.nochim_DENG7)) 

taxa_DENG7 <- read.csv("Deng_taxa_16S.csv", row.names = 1)
row.names(taxa_DENG7) = colnames(seqtab.nochim_DENG7)
taxa_DENG7 <- as.matrix(taxa_DENG7)

metadata_DENG7 <- read.csv("Metadata_DENG_2017.csv", stringsAsFactors = TRUE, row.names = 1)

ps_DENG7 <- phyloseq(
  otu_table(seqtab.nochim_DENG7, taxa_are_rows=FALSE),
  sample_data(metadata_DENG7),
  tax_table(taxa_DENG7))

ps_DENG7
```


# Importing files, renaming ASVs and creating phyloseq object - DENG_2018 
```{r importing files deng8}
seqtab.nochim_DENG8 <- readr::read_csv("Deng_2_seqtab.nochim_16S.csv")
seqtab.nochim_DENG8 <- as.data.frame(seqtab.nochim_DENG8)
rownames(seqtab.nochim_DENG8) <- seqtab.nochim_DENG8[, 1]
seqtab.nochim_DENG8 <- seqtab.nochim_DENG8[, -1]
colnames(seqtab.nochim_DENG8) = paste0("ASV", 1:ncol(seqtab.nochim_DENG8)) 

taxa_DENG8 <- read.csv("Deng_2_taxa_16S.csv", row.names = 1)
row.names(taxa_DENG8) = colnames(seqtab.nochim_DENG8)
taxa_DENG8 <- as.matrix(taxa_DENG8)

metadata_DENG8 <- read.csv("Metadata_DENG_2018.csv", stringsAsFactors = TRUE, row.names = 1)

ps_DENG8 <- phyloseq(
  otu_table(seqtab.nochim_DENG8, taxa_are_rows=FALSE),
  sample_data(metadata_DENG8),
  tax_table(taxa_DENG8))

ps_DENG8
```

# Importing files, renaming ASVs and creating phyloseq object - FENG 
```{r importing files feng}
seqtab.nochim_FENG <- readr::read_csv("Feng_seqtab.nochim_16S.csv")
seqtab.nochim_FENG <- as.data.frame(seqtab.nochim_FENG)
rownames(seqtab.nochim_FENG) <- seqtab.nochim_FENG[, 1]
seqtab.nochim_FENG <- seqtab.nochim_FENG[, -1]
colnames(seqtab.nochim_FENG) = paste0("ASV", 1:ncol(seqtab.nochim_FENG)) 

taxa_FENG <- read.csv("Feng_taxa_16S.csv", row.names = 1)
row.names(taxa_FENG) = colnames(seqtab.nochim_FENG)
taxa_FENG <- as.matrix(taxa_FENG)

metadata_FENG <- read.csv("Metadata_FENG.csv", stringsAsFactors = TRUE, row.names = 1)

ps_FENG <- phyloseq(
  otu_table(seqtab.nochim_FENG, taxa_are_rows=FALSE),
  sample_data(metadata_FENG),
  tax_table(taxa_FENG))

ps_FENG
```

# Importing files, renaming ASVs and creating phyloseq object - Current work 
```{r importing files current work}
seqtab.nochim <- readr::read_csv("seqtab.nochim.csv")
seqtab.nochim <- as.data.frame(seqtab.nochim)
rownames(seqtab.nochim) <- seqtab.nochim[, 1]
seqtab.nochim <- seqtab.nochim[, -1]
colnames(seqtab.nochim) = paste0("ASV", 1:ncol(seqtab.nochim)) 

taxa <- read.csv("taxa.csv", row.names = 1)
row.names(taxa) = colnames(seqtab.nochim)
taxa <- as.matrix(taxa)

metadata <- read.csv("metadata.csv", stringsAsFactors = TRUE, row.names = 1)

ps <- phyloseq(
  otu_table(seqtab.nochim, taxa_are_rows=FALSE),
  sample_data(metadata),
  tax_table(taxa))

# excluding foam samples that were not used in this work
ps_Without_FOAM <- subset_samples(ps, Phase != "R1_Foam")
ps_Without_FOAM <- subset_samples(ps_Without_FOAM, Phase != "R2_Foam")
ps_Without_FOAM <- subset_samples(ps_Without_FOAM, Phase != "R3_Foam")
ps_Without_FOAM
```

# Joining phyloseq objects
```{r building phylo}
ps_all <- merge_phyloseq(ps_Without_FOAM, ps_auer, ps_DENG7, ps_DENG8, ps_FENG)
ps_all
```

# Rarefying
```{r rarefying}
set.seed(1)

counts <- sample_sums(ps_all)
sort(counts)

ps_all_rar<- rarefy_even_depth(ps_all, sample.size = 11738,
                  rngseed = FALSE, replace = FALSE, trimOTUs = TRUE, verbose = TRUE)
ps_all_rar
```

# Permanova
```{r permanova}
dist <- phyloseq::distance(ps_all_rar, method="bray")
head(dist)

metadata_dist <- data.frame(sample_data(ps_all_rar))
head(metadata_dist)

test.adonis_studies <- adonis2(dist ~ Paper, data = metadata_dist)
test.adonis_studies
```

# Taxonomy plot - Family
```{r tax plot family}
custom_col42 = c("#781156","#A51876","#D21E96","#E43FAD","#EA6CC0","#F098D3","#114578","#185EA5","#1E78D2","#3F91E4","#6CABEA","#98C4F0","#117878","#18A5A5","#3FE4E4","#6CEAEA","#98F0F0", "#117845","#18A55E","#1ED278","#3FE491","#6CEAAB","#98F0C4","#787811","#A5A518","#D2D21E","#E4E43F","#EAEA6C","#F0F098","#F7F7C5","#784511","#A55E18","#D2781E","#E4913F","#EAAB6C","#F0C498","#781122","#A5182F","#D21E2C","#E43F5B","#EA6C81","#F098A7")

y1p <- tax_glom(ps_all_rar, taxrank = 'Family') # agglomerate taxa
y3p <- transform_sample_counts(y1p, function(x) x/sum(x)) #get abundance in %
y4p <- psmelt(y3p) # create dataframe from phyloseq object
y4p$Family <- as.character(y4p$Family) #convert to character
y4p$Family[y4p$Abundance < 0.05] <- " Family < 5% abund." #rename genera with < 1% abundance

#set color palette to accommodate the number of genera
colourCount <- length(unique(y4p$Family))
getPalette = colorRampPalette(custom_col42, interpolate="spline")

# all samples
#png("figures_article/Tax_Family_days.png", width = 10, height = 6, units = "in", res=300)
p_p <- ggplot(data=y4p, aes(x=Sample, y=Abundance, fill=Family))
p_p + geom_bar(aes(), stat="identity", position="stack") + 
  facet_wrap(~ Paper , scales = "free", labeller = labeller(.multi_line = FALSE)) +
  scale_fill_manual(values=getPalette(colourCount)) + 
  xlab("Samples") + ylab("Relative Abundance") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), legend.position="bottom")
#dev.off()

```
